---
title: "R Notebook"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

#Check working directory and clean workspace
```{r}
getwd()

rm(list=ls())
```
#Importing dataset
```{r}
#import ManyBabies1DataExport(05-01-2018) wide dataset
wide_dataset <- read.delim("../data/ManyBabiesDataExport.csv",sep = ",")
trial_data <- read.delim("../data/MB1_TrialDataUpdated.csv", sep = ",", na.strings = c("NA", ""),stringsAsFactors = F)
orders <- read.delim("../data/MB1_orders.csv", sep = ",")
participants <- read.delim("../data/MB1_MasterSubjectList.csv", sep = ",")
participant_template <- read.delim("../data/ManyBabies Individual Lab Participants Data.csv", sep = ",", header = T)

```

```{r}
#loading  tidyverse packages
library(tidyverse)
```
#Cleaning participant data
```{r}
#inputting subject ID from MSL with their orders and keeping only those 2 columns of information
participants2 <- participants %>%
  mutate(subid = paste(MB_group_6_vs_12, subid..ManyBabies., sep = "_")) %>%
  mutate(order = trial_order_manybabies) %>%
  select(subid, order, lab)
```

#Manipulating Tobii data
```{r}
#renaming participant column (X) name variable to subid
wide_dataset <- rename(wide_dataset,   subid = X)

#converting from wide to long dataset and renaming variables in long dataset
long_dataset = wide_dataset %>%
 gather(key =trial, value = total_fixation, Total.Fixation.Duration..Include.Zeros._ADS1_ADS1_Sum:Total.Fixation.Duration..Include.Zeros._train2_train2_Sum.3, factor_key=TRUE) %>%
  arrange(subid, trial) %>%
  mutate(trial = gsub("Total.Fixation.Duration..Include.Zeros._", "", trial)) %>%
  mutate(trial = sub("_.*", "", trial)) 

#removing the "all recordings" rows
long_dataset <- subset(long_dataset, subid != "All Recordings")

#creating new variable column for trial type
long_dataset$trial_type <- substr(long_dataset$trial, 1,3)

#moving columns around 
long_dataset <- long_dataset [,c(1, 4, 2, 3)]

#finding the largest value within every trial (signifying the order participated it)
long_dataset_clean <- long_dataset %>%
  dplyr::group_by(subid, trial) %>%
  dplyr::summarize(total_fixation = max(total_fixation, na.rm=T)) 

#creating new variable column for trial type in new dataframe
long_dataset_clean$trial_type <- substr(long_dataset_clean$trial, 1,3)

#seperating _ into its own columns
long_dataset_clean <- separate(long_dataset_clean, subid, into = c("STUDY", "STUDYID", "WORSE"))

#deleting columns "BAD" and "WORSE"
long_dataset_clean[[3]] <- NULL

#merging BABY and BAD column together to create subid column
long_dataset_clean <- transform(long_dataset_clean, subid=paste(STUDY, STUDYID, sep="_"))

#deleting individual columns "BABY" and "BAD"
long_dataset_clean[[1]] <- NULL
long_dataset_clean[[1]] <- NULL

#reordering columns
long_dataset_clean <- long_dataset_clean [,c(4, 3, 1, 2)]

#renaming values of "tra" (in trial_type columns) to "training"
long_dataset_clean$trial_type[long_dataset_clean$trial_type == "tra"] <- "training"



```
#cleaning MSL to remove all GazeFollowing-specific columns
```{r}
participants2 <- participants2[, ]
```

#Cleaning trial data
```{r}
trial_data <- trial_data %>%
  mutate(subid = as.factor(subid)) %>% 
  mutate(trial_type = as.factor(trial_type)) %>% 
  mutate(trial_num = as.numeric(trial_num)) %>%
    mutate(total_trial_time = as.numeric(total_trial_time)) %>%
  select(subid, trial_num, total_trial_time, trial_error, trial_error_type, notes) 

```

#Merging Tobii with participants and order and cleaning up file
```{r}
joined_dataset_clean_temp <- long_dataset_clean %>%
  left_join(participants2, by ="subid") %>% # get orders for each participant
  left_join(orders) %>%
  rename(looking_time = total_fixation)


joined_dataset_clean <- joined_dataset_clean_temp %>% # gets trial orders
  left_join(trial_data) %>%
  mutate(total_trial_time = as.numeric(total_trial_time)) %>%
  rename(stimulus = trial) %>% 
  select(lab, subid, trial_type, stimulus, trial_num, looking_time, total_trial_time, trial_error, trial_error_type, notes) %>%
  mutate(trial_num =ifelse(trial_type =="training",trial_num-3, trial_num -2)) %>% #So first test trial is trial 1, and training trials are -2 and -1
  arrange(subid, trial_num)

joined_dataset_clean$looking_time[is.na(joined_dataset_clean$total_trial_time)] <- NA # No looking with no trial time should be NAs

#Replace NA values necessary with NC
replace_na(joined_dataset_clean, list(looking_time = "NC"))
replace_na(joined_dataset_clean, list(total_trial_time = "NC"))

#converting all NAs in lab column to irl-concordia
joined_dataset_clean <- replace_na(joined_dataset_clean, list(lab = "irl-concordia"))
```
# data checking to verify that all of 0 looking time are marked as errors
```{r}
joined_dataset_clean %>% filter(looking_time == 0 & trial_error == "noerror") %>% write.csv("../data/MB1_checkingerrors.csv" )
```

# Exporting just long dataset to type trial lengths into
```{r}
write.csv(joined_dataset_clean_temp, "../data/long_dataset_clean.csv")
```

#Exporting R table to .csv file
```{r}
write.csv(joined_dataset_clean, "../data/MB1_TrialData.csv", row.names = FALSE)
```

#Creating final participants file

```{r}

mb_columns <- colnames(participant_template)[1:49]

participants_mono <- participants %>%
  filter(lang_group  == "monolingual") %>%
  filter(!is.na(trial_order_manybabies)) %>%
  unite(subid, MB_group_6_vs_12, subid..ManyBabies., sep = "_", remove = T) %>%
  select(-BABYID, -do.birth, -do.participation, -subid..GazeFollowing., -GF_group_6_vs_12, -trial_order_gazefollowing, -session_error_gazefollowing, -session_error_type_gazefollowing, -second_session_GazeFollowing) %>%
  rename(trial_order = trial_order_manybabies) %>%
  rename(session_error = session_error_manybabies) %>%
  rename(session_error_type = session_error_type_manybabies) %>%
  rename(second_session = second_session_ManyBabies) %>%
  rename(race_ethinicty = race.ethnicity) %>%
  rename(curr_earinfection = current_earinfection) %>%
  mutate(notes = recode(Notes, "post-undergrad" = "RA is post-undergrad")) %>%
  select(mb_columns) %>% # see https://gist.github.com/djhocking/62c76e63543ba9e94ebe
  arrange(subid)
  # 
  # 
  # 
  # select(lab, subid, method, RA, age_days, trial_order, session_error, session_error_type, notes, second_session, lang_group, lang1, lang1_exposure, lang2, lang2_exposure, lang3, lang3_exposure, lang4, lang4_exposure, caregiver_seat, MacArthur_Ladder, participant_gender, race_ethnicity, parentA_gender)


```

# Write final monolingual participants to file
```{r}

write.csv(participants_mono, "../data/irl-concordia_Monolingual_Participant_Info.csv")


```


# Creating final trials data
```{r}


final_dataset_monolinguals <- joined_dataset_clean[joined_dataset_clean$subid %in% participants_mono$subid, ]
```

# Write final monolingual dataset to file
```{r}

write.csv(final_dataset_monolinguals, "../data/irl-concordia_Monolingual_Trial_Data.csv")


```


